{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-02-15 landslide-lecture/2018/lecture.tex
landslide lecture - copy 2017 2018

\title[Landslide]{{\bf Landslide: A New Race-Finding Tool for 15-410}}
\subtitle[]{ {\em more clever than } ``\texttt{mandelbrot}'' {\em since 2011.}}
\author[Ben Blum]{Ben Blum \texttt{(bblum@andrew.cmu.edu)}}

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2017, February 20}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing race conditions in a new way}
	\begin{itemize}
		\item Case study (example)
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Research Technique: ``Systematic testing''}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item Automatically choosing preemption points
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

\begin{frame}{About The Project}
	\textbf{About me}: 6th year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		%\item 1st graduate year was 5th year M.S., rest in Ph.D. program
		\item Landslide's publication history
		\begin{itemize}
			\item Master's thesis
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
			\end{itemize}
			\item Conference paper (OOPSLA 2016)
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/oopsla.pdf}
			\end{itemize}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Iterative Deepening of Preemption Points}
	Goal: Run the best tests for a given CPU budget.
	%\begin{itemize}
	%	\item Technique: ``Iterative Deepening''
	%\end{itemize}
	\linegap

	Based on experience from past 15-410 student volunteers
	\begin{itemize}
		%\item Students worked best with an iterative process
		\item ``Start small, then add more preemption points as time allows''
		\item Landslide now automates this process
	\end{itemize}
	\linegap

	Named after analogous technique in chess AI.
	\begin{itemize}
		\item Chess search is DFS limited by max number of moves (ply).
		\item Chess AIs repeat DFS, increasing ply, until timeout.
	\end{itemize}
\end{frame}

@@

@@

\begin{frame}{Previous Semesters}
	%Spring 2015: 7 groups signed up to use Landslide; 5 found bugs
	%%\begin{itemize}
	%%	\item (that they didn't find during stress testing)
	%%\end{itemize}
	%%%%% hopefully!
	%\linegap

	%Among all groups, 38 total tests were run
	%\begin{itemize}
	%	\item 21 tests ran to completion without finding bugs
	%	\item 4 {\em deterministic} bugs
	%	\item 13 distinct {\em non-deterministic} bugs
	%		\begin{itemize}
	%			\item 6 in {\tt thr\_exit\_join} \small (use-after-free, deadlock, NULL, infinite loop...)
	%			\item 4 in {\tt paraguay} (deadlock, infinite loop)
	%			\item 2 in {\tt broadcast\_test} (deadlock, infinite loop)
	%			\item 1 in {\tt rwlock\_downgrade\_read\_test} (use-after-free)
	%		\end{itemize}
	%		%\begin{itemize}
	%		%	\item 6 of which required data-race PPs to expose
	%		%\end{itemize}
	%\end{itemize}
	S'15, F'15, S'16: 48 groups signed up to use Landslide; 38 found bugs
	% 18 groups found races
	%\begin{itemize}
	%	\item (that they didn't find during stress testing)
	%\end{itemize}
	%%%% hopefully!
	\linegap

	44 {\em deterministic} bugs (e.g. swexn, initialization)
	\linegap

	85 distinct {\em non-deterministic} bugs (among 32 groups)
	\begin{itemize}
		\item 24 groups (75\%) fixed $\ge 1$ such bug
			\begin{itemize}
				\item (as verified by running Landslide again -- not a guarantee!)
			\end{itemize}
		\item 15 groups (47\%) fixed {\em all} such bugs
		\item Most ambitious group: 11 distinct races found + fixed!
	\end{itemize}
	\linegap
	% 44 dets 85 races

	%Among all groups, 185 distinct tests were run
	%\begin{itemize}
	%	\item 137 tests ran to completion without finding bugs
	%	\item 21 {\em deterministic} bugs
	%\end{itemize}
	%\pause % ask: "who can tell me why landslide might find a *deterministic* bug not found during past testing? A: use-after-free
	%\linegap

	%38 distinct {\em non-deterministic} bugs
	%\begin{itemize}
	%		%\begin{itemize}
	%		%	\item 7 of which required data-race PPs to expose
	%		%\end{itemize}
	%	\item 11 groups (50\%) fixed all their bugs
	%		\begin{itemize}
	%			\item (as verified by running Landslide again -- not a guarantee!)
	%		\end{itemize}
	%	\item 2.7 bugs ``fixed'' (see above) on average per group
	%	\item Most ambitious group: 11 distinct races found + fixed!
	%\end{itemize}
\end{frame}

@@

	To participate\dots
	\begin{itemize}
		%\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item \url{http://tinyurl.com/landslide-p2-s17}
			\end{itemize}
		\item Optional ``Landslide clinic'' for in-person tech (or moral) support
			\begin{itemize}
				\item Later this week, room and time TBD
			\end{itemize}
		%\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	\linegap

	{\em Really} interested?
	\begin{itemize}
		\item I know of an M.S. thesis topic if you want to do post-OS research.
	\end{itemize}
\end{frame}

>>>>

\title[Landslide]{{\bf Landslide: A New Race-Finding Tool for 15-410}}
\subtitle[]{ {\em more clever than } ``\texttt{agility\_drill}'' {\em since 2011.}}
\author[Ben Blum]{Ben Blum \texttt{(bblum@andrew.cmu.edu)}}

\institute[CMU 15-410]{Carnegie Mellon University - 15-410}
\date[]{2017, September 27}

@@

\begin{frame}{Outline}
	\textbf{Theory: Seeing concurrency bugs in a new way}
	\begin{itemize}
		\item Case study (example)
		\item Tabular execution traces
		\item The execution tree
	\end{itemize}
	{\bf Research Technique: ``Systematic testing''}
	% Say: A way of enumerating all possible interleavings of a concurrent test case.
	\begin{itemize}
		\item Preemption points
		\item Challenges and feasibility
	\end{itemize}
	{\bf Tool: Landslide}
	\begin{itemize}
		\item How it works
		\item Automatically choosing preemption points
		\item User study (that's you!)
			% !!! SAY !!! : If you are struggling, this is not for you, etc.
	\end{itemize}
\end{frame}

@@

\begin{frame}{About The Project}
	\textbf{About me}: Final(?) year graduate student, advised by Garth Gibson
	\begin{itemize}
		\item TAed 15-410 for 3 semesters during undergrad
		%\item 1st graduate year was 5th year M.S., rest in Ph.D. program
		\item Landslide's publication history
		\begin{itemize}
			\item Master's thesis
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/thesis.pdf}
			\end{itemize}
			\item Conference paper (OOPSLA 2016)
			\begin{itemize}
				\item \url{http://www.contrib.andrew.cmu.edu/~bblum/oopsla.pdf}
			\end{itemize}
		\end{itemize}
	\end{itemize}
	\pause
	\linegap

@@

%\begin{frame}{Iterative Deepening of Preemption Points}
%	Goal: Run the best tests for a given CPU budget.
%	%\begin{itemize}
%	%	\item Technique: ``Iterative Deepening''
%	%\end{itemize}
%	\linegap
%
%	Based on experience from past 15-410 student volunteers
%	\begin{itemize}
%		%\item Students worked best with an iterative process
%		\item ``Start small, then add more preemption points as time allows''
%		\item Landslide now automates this process
%	\end{itemize}
%	\linegap
%
%	Named after analogous technique in chess AI.
%	\begin{itemize}
%		\item Chess search is DFS limited by max number of moves (ply).
%		\item Chess AIs repeat DFS, increasing ply, until timeout.
%	\end{itemize}
%\end{frame}

@@

\begin{frame}{What makes a Landslide-friendly test}
	``Why not {\tt largetest}, {\tt juggle}, {\tt cyclone}, {\tt agility\_drill}...?''
	\linegap

	Sample code from {\tt cyclone}: % abbreviated
	\begin{center}
	\begin{tabular}{l}
		\texttt{for (i = 0; i < MAX\_MISBEHAVE /* 64 */; i++) \{} \\
		\texttt{~~~~misbehave(i);} \\
		\texttt{~~~~tid = thr\_create(child\_fn, i);} \\
		\texttt{~~~~thr\_join(tid, \&status);} \\
		\texttt{\}}
	\end{tabular}
	\end{center}
	\pause
	\linegap

	Stress tests expose various interleavings using big loops/many threads;
	\linegap

	Landslide finds many interleavings itself; even if the test case has no loops.
	\vspace{0.01in}

	How long do you think Landslide would take to test {\tt cyclone}..?
\end{frame}

@@

\begin{frame}{Previous Semesters}
	%% S'15, F'15, S'16: 48 groups signed up to use Landslide; 38 found bugs
	S'15, F'15, S'16, F'16, S'17: 79 groups signed up to use Landslide; 59 found bugs
	\linegap

	%%44 {\em deterministic} bugs (e.g. swexn, initialization)
	91 {\em deterministic} bugs (e.g. swexn, initialization)
	\linegap

	%%85 distinct {\em non-deterministic} bugs (among 32 groups)
	113 distinct {\em non-deterministic} bugs (among 44 groups)
	\begin{itemize}
		\item 33 groups (75\%) fixed $\ge 1$ such bug
		%%\item 24 groups (75\%) fixed $\ge 1$ such bug
			\begin{itemize}
				\item (as verified by running Landslide again -- not a guarantee!)
			\end{itemize}
		\item 24 groups (55\%) fixed {\em all} such bugs
		%%\item 15 groups (47\%) fixed {\em all} such bugs
		\item Most ambitious group: 11 distinct races found + fixed!
	\end{itemize}
	\linegap
\end{frame}

@@

	To participate\dots
	\begin{itemize}
		%\item Review this lecture and study information sheet
		\item Meet prerequisites of passing P2 tests
		\item Complete sign-up form online to get further instructions
			\begin{itemize}
				\item (watch your email for the link)
				%\item \url{http://tinyurl.com/landslide-p2-s17}
			\end{itemize}
		\item Optional ``Landslide clinic'' for in-person tech (or moral) support
			\begin{itemize}
				\item Next week, room and time TBD
			\end{itemize}
		%\item Use Landslide in addition to stress tests until P2 is due!
	\end{itemize}
	%% im out of time :<<
	%\linegap

	%{\em Really} interested?
	%\begin{itemize}
	%	\item I know of an M.S. thesis topic if you want to do post-OS research.
	%\end{itemize}
\end{frame}

### 2018-02-15 thesis/landslide.tex
landslide lecture - copy 2017 2018

I use the {\sc Djit+} rules for reads and writes rather than the {\tt FastTrack} ones,
even though they incur $O(n)$ runtime in the size of the vector clocks more often:
because Landslide tests should be limited to few threads to manage the state space size,
$O(n)$ is usually in the single digits, so I optimize for code simplicity.

\begin{itemize}
	\item Reads and Writes ({\tt memory.c})
	\begin{itemize}
		\item \textsc{Djit+ read/write same epoch} - {\tt vc\_eq()} case of {\tt add\_lockset\_to\_shm()}
		\item \textsc{Djit+ read/write} - {\tt vc\_happens\_before()} case of {\tt check\_locksets()}
	\end{itemize}
	\item Synchronization ({\tt schedule.c})
	\begin{itemize}
		\item \textsc{FT acquire}
			\begin{itemize}
				\item {\tt kern\_mutex\_\{,try\}locking\_done()} cases of {\tt kern\_update\_state\_\allowbreak{}machine()}
				\item {\tt user\_mutex\_\{,try\}lock\_exiting()} cases of {\tt user\_update\_state\_\allowbreak{}machine()}
				\item {\tt cli} case of {\tt kern\_update\_state\_machine()} (Pintos only)
				\item {\tt cli}/{\tt sti} lock handoff case in {\tt sched\_update()} (Pebbles only)
			\end{itemize}
		\item \textsc{FT release}
			\begin{itemize}
				\item {\tt kern\_mutex\_unlocking()} case of {\tt kern\_update\_state\_machine()}
				\item {\tt user\_mutex\_unlock\_entering()} case of {\tt user\_update\_state\_machine()}
				\item {\tt sti} case of {\tt kern\_update\_state\_machine()} (Pintos only)
				\item {\tt cli}/{\tt sti} lock handoff case in {\tt sched\_update()} (Pebbles only)
			\end{itemize}
		\item \textsc{FT fork} - {\tt agent\_fork()}
		\item \textsc{FT join} - {\tt sched\_unblock()} case of {\tt kern\_update\_state\_machine()} (Pebbles only)
	\end{itemize}
\end{itemize}

\subsection{Iterative Context Bounding}
\label{sec:landslide-icb}
\subsection{Heuristic loop, deadlock, and synchronization detection}

>>>>

I use the {\sc Djit+} rules for reads and writes rather than the {\sc FastTrack} ones,
even though they more often incur $O(n)$ runtime in the size of the vector clocks:
because Landslide tests should be limited to few threads in order to manage the state space size,
$n$ is always in the single digits, so I optimize for code simplicity.

\begin{itemize}
	\item Reads and writes ({\tt memory.c})
	\begin{itemize}
		\item \textsc{Djit+ read/write same epoch} - {\tt vc\_eq()} case of {\tt add\_lockset\_to\_shm()}
		\item \textsc{Djit+ read/write} - {\tt vc\_happens\_before()} case of {\tt check\_locksets()}
	\end{itemize}
	\item Synchronization ({\tt schedule.c})
	\begin{itemize}
		\item \textsc{FT acquire}
			\begin{itemize}
				\item {\tt kern\_mutex\_\{,try\}locking\_done()} cases of {\tt kern\_update\_state\_\allowbreak{}machine()}
				\item {\tt user\_mutex\_\{,try\}lock\_exiting()} cases of {\tt user\_update\_state\_\allowbreak{}machine()}
				\item {\tt cli} case of {\tt kern\_update\_state\_machine()} (Pintos only)
				\item {\tt cli}/{\tt sti} lock handoff case in {\tt sched\_update()} (Pebbles only)
			\end{itemize}
		\item \textsc{FT release}
			\begin{itemize}
				\item {\tt kern\_mutex\_unlocking()} case of {\tt kern\_update\_state\_machine()}
				\item {\tt user\_mutex\_unlock\_entering()} case of {\tt user\_update\_state\_machine()}
				\item {\tt sti} case of {\tt kern\_update\_state\_machine()} (Pintos only)
				\item {\tt cli}/{\tt sti} lock handoff case in {\tt sched\_update()} (Pebbles only)
			\end{itemize}
		\item \textsc{FT fork} - {\tt agent\_fork()}
		\item \textsc{FT join} - {\tt sched\_unblock()} case of {\tt kern\_update\_state\_machine()}
			(Pebbles only; Pintos case is handled by above {\tt cli}/{\tt sti} cases in context switch)
	\end{itemize}
\end{itemize}

\subsection{Iterative Context Bounding}
\label{sec:landslide-icb}

Iterative Context Bounding \cite{chess-icb} is a state space exploration strategy
that prioritizes interleavings with fewer total preemptions first.
Let $P(I)$ denote the number of preemptions in an interleaving $I$.
Then, to summarize in pseudocode a na\"ive exploration of some state space $S$ as:

\begin{algorithm}[h]
	\ForEach{$I \in S$}{
		Execute($I$)
	}
\end{algorithm}

ICB's approach could likewise be summarized as follows:

\begin{algorithm}[h]
	\For{$n \in [0..\mathsf{max}_P(S)]$}{
		\ForEach{$I \in S, P(I) \le n$}{
			Execute($I$)
		}
	}
\end{algorithm}

%% stupid imperative way to write it
%\begin{algorithm}[h]
%	\For{$n \in [0..\infty]$}{
%		$\mathsf{skipped\_any} := \mathsf{false}$ \\
%		\ForEach{$I \in S$}{
%			\uIf{$P(I) \le n$}{
%				Execute($I$)
%			} \Else {
%				$\mathsf{skipped\_any} := \mathsf{true}$
%			}
%		}
%		\If{$\neg \mathsf{skipped\_any}$}{
%			{\bf Break}
%		}
%	}
%\end{algorithm}

However, when combined with DPOR to prune equivalent interleavings,
DPOR's reduction might not be sound with respect to the subset of $S$ under $P \le n$.
Prior work introduced Bounded Partial Order Reduction (BPOR) \cite{bpor},
a compatibility extension to DPOR for ICB to address this problem.
I refer the reader interested in the algorithm's details to the citation;
this section will discuss Landslide's implementation.

% TODO: say how u do it

\subsection{Heuristic loop, deadlock, and synchronization detection}
